---
title: "STAT 370: Theoretical Statistics - Problem Set 1"
author: "Group 5: Bonnie Lin, Daniel Njoo, Will Zaubler"
date: "January 25, 2018"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

```{r}
library(dplyr)
set.seed(370)
numsim <- 10000
```


## Question 3

A machine produces bearings with mean diameter 3.00 inches and standard deviation 0.01
inches. Bearings with diameters in excess of 3.02 inches or less than 2.98 inches will be deemed
to fail to meet quality specifications. Assume that the underlying process is normally distributed.
Approximately what fraction of this machine’s production will fail to meet specifications? Simulate
from a normal distribution to check your analytic answer. Then assume that the underlying
process is Gamma distributed with the same mean and sd. Simulate from an appropriate Gamma
distribution to check your analytic answer. How important is the normality assumption?

### Analytical Answer

Assuming Normal distribution:

Let b represent a bearing produced by the machine. Then:

$P(b>3.02$ or $b<2.98)$ 

=$(1-P(b\leq3.02))+P(b\geq2.98)$ — since it's a continuous probability distribution

=(1-pnorm(3.02,3,0.01))+pnorm(2.98,3,0.01)

=`r (1-pnorm(3.02,3,0.01))+pnorm(2.98,3,0.01)`

--

Assuming Gamma distribution

Let b represent a bearing produced by the machine. Then:

$E[X]=\frac{\alpha}{\beta}$ and $V[X]\frac{\alpha}{\beta^2}$, so solving for $\alpha$ and $\beta$, we find $\alpha=90000$ and $\beta=30000$, so 

$(1-P(b\leq3.02))+P(b\geq2.98)$

=(1-pgamma(3.02,90000,30000))+pgamma(2.98,90000,30000)

=`r (1-pgamma(3.02,90000,30000))+pgamma(2.98,90000,30000)`

### Empirical Answer 

#### Normal 

We create a vector of 10,000 numbers that follow a Normal Distribution and determine the number of those within the vector that fulfill the conditions, x>3.02 or x<2.98. Then, we found the proportion of these out of 10,000.

```{r}
norm<-rnorm(numsim, 3, 0.01)
sum(norm>3.02 | norm<2.98)/numsim
```

Sampling under a normal distribution, we find that 0.0447 of the bearings fail to meet the specification. We confirm this with a visualization that shows that 0.046 of the distribution falls within our specified conditions.

```{r}
xpnorm(c(2.98,3.02),3,0.01,plot=T)
```

#### Gamma 

We create a vector of 10,000 numbers that follow a Gamma Distribution and determine the sum of those within the vector that fulfill the conditions, x>3.02 or x<2.98. Then, we found the proportion of these out of 10,000.

```{r}
gamma<-rgamma(n=numsim, shape=90000, rate=30000)
sum(gamma>3.02 | gamma<2.98)/numsim
```

We can see that the simulation gives a good estimate of the empirical answer (0.05). A visualization shows us that 0.046 (just like the Normal) fulfills the conditions.

```{r}
xpgamma(c(2.98,3.02),shape=90000,rate=30000,plot=T)
```


## Question 5

A personnel manager for a certain industry has records of the number of employees absent
per day. The average number absent is 5.5, and the standard deviation is 2.5. Because there are
many days with zero, one or two absent and only a few with more than 10 absent, the frequency
distribution is highly skewed. The manager wants to publish an interval in which at least 75% of
these values lie. Use the result from the previous problem to find such an interval. Generate data
from a distribution that matches these parameters and compare your result to the true value.

### Analytical Answer 

Given Chebyshev's Inequality, we know that $P(|X-\mu|\geq k\sigma)\geq 1-\frac{1}{k^2}$, and that $1-\frac{1}{k^2}=0.75$ so $k=2$.

So the interval is $\mu\pm2\sigma=(0.5,10.5)$

### Empirical Answer 

We want a pdf that has the highest probability at P(A=0), where A is the number of employees absent, and decreases rapidly such that P(A>=10) is marginal. We know that the parameters of the pdf are mean 5.5, and sd 2.5.

So we created a gamma distribution using parameters $r=4.84$, and $\lambda=0.88$, which we solved for from the formulae for E(X) and V(X) of a gamma distribution. 

We then created 10000 values from this distribution, and checked that the mean and sd of these values equals those of the distribution.

Finally, we find the 75% interval via finding the 12.5% and 87.5% tiles of our gamma distribution and find the interval (2.84, 8.4).

```{r}
x <- rgamma(n=numsim, shape=4.84, rate=.88) ; mean(x) ; sd(x)

qgamma(c(.125, .875), shape=4.84, rate=.88)
```

Using our empirical approach we therefore found an interval of (2.84, 8.40), which is much narrower than our analytical approach via Chebyshev's which created an interval of (0.5, 10.5) — extremely conservative! A graphical comparison is shown below:

```{r}
xpgamma(c(2.84, 8.4), shape=4.84, rate=.88, return="plot") + geom_vline(xintercept=c(0.5,10.5))
```


## Question 9

Shuffle an ordinary deck of 52 playing cards containing four aces. Then turn up cards from
the top until the first ace appears. On the average, how many cards are required to produce the
first ace?

### Analytical Answer 

4 Aces can divide a deck into 5 portions. Each of these portions would on average have $\frac{48}{5}=9.6$ cards in them.

Therefore the average number of cards required to draw the first ace would be $9.6+1=10.6$, in order to include the drawing of the ace itself.

### Empirical Answer 

We create our deck here, with 4 Aces and 48 other 'no's.

```{r}
deck <- c(rep("Ace", 4), rep("no", 48))
```

We then shuffle the deck; here we show the first 6 cards drawn.

```{r}
set.seed(9)
shuffled.deck <- shuffle(deck); head(shuffled.deck) # Example
```

We then search through all 52 cards for the locations of the Aces, which are given to as location indices in the 52-long vector. 

```{r}
grep("Ace", shuffled.deck) # Location of Aces
```

If we min() this we can find the location of the first Ace of drawn.

```{r}
grep("Ace", shuffled.deck) %>%  min()
```

We then incorporate this into a function which we run 10000 times with do().

```{r}
question9 <- function() {

  shuffled.deck <- shuffle(deck)
  locations <- grep("Ace", shuffled.deck)
  num.before <- min(locations)
  return(num.before)

}

num.cards.for.ace <- do(numsim) * question9()
num.cards.for.ace %>% 
  unlist() %>% 
  mean()  # %>% means passing previous variable as parameter of next function; think "then"
```

And we find a mean of 10.46, with a 95% CI of (10.30, 10.62), which contains our analytic answer of 10.6

```{r}
t.test(~ question9, data = num.cards.for.ace)
```

A histogram of the cards drawn from 10000 simulations is  shown below:

```{r}
as.numeric(unlist(num.cards.for.ace)) %>% hist(main="Histogram of cards drawn till ace from shuffled deck")
```

